### 2021-04-09 ch06_paper_gpp/paper_gpp.tex
Fixed capitalization for step XY

Here we briefly describe the data and methods used in this study.
\Cref{subsec:06:step_1} reviews the method of a previously published emergent
constraint on \ac{CO2} fertilization from \textcite{Wenzel2016} that we apply
here to our study to rescale the climate model output (step 1). This includes a
discussion of the caveats that are associated with this emergent constraint
that attributes changes in the observed increase in the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} entirely to the increase in atmospheric \ac{CO2}
concentration, yet other studies provide indications that part of this increase
is due to climate and land use changes \autocite{Bastos2019, Forkel2016,
  Piao2018, Zhao2016}. In \cref{subsec:06:step_2} the \ac{ML} technique used in
step 2 to constrain gridded \ac{GPP} projections is presented.
\Cref{fig:06:schematic_steps} shows an overview of our two-step approach.
\Cref{subsec:06:uncertainty_and_skill} describes how uncertainty and predictive
skill are quantified. A complete list of all \acs{CMIP}5 models used in this
study is given in \cref{tab:app:b:cmip5_models}. To apply the emergent
constraint in step 1, we need emission-driven \acs{CMIP}5 simulations. The
number of \acp{ESM} included is therefore seven, as in \textcite{Wenzel2016}.
More details on the methods and the experimental setup are given in
\cref{sec:app:si_for_paper_gpp}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{ch06_paper_gpp/figs/1.pdf}
  \caption[
    Schematic illustration of our two-step approach.
  ]{
    Schematic illustration of our two-step approach. In step 1 an emergent
    constraint by \textcite{Wenzel2016} is used to constrain the global mean
    fractional change in \acf{GPP} over the \nth{21} century. Moreover, this
    constraint is used to rescale two different gridded target variables:
    absolute \acs{GPP} at the end of the \nth{21} century and fractional
    \acs{GPP} change over the \nth{21} century. In step 2, a \acl{ML} model is
    used to constrain these two target variables (step 2a: absolute \acs{GPP};
    step 2b: fractional \acs{GPP} change) in two independent approaches.
    \AdaptedFrom{Schlund2020}.
  }
  \label{fig:06:schematic_steps}
\end{figure}

@@

Therefore, our target variable in step 1 is the global mean fractional
\acs{GPP} change over the \nth{21} century $f$ in the \acs{RCP}8.5 scenario,
which is calculated from the 10-year mean \ac{GPP} at the end of the \nth{21}
century in the emission-driven fully coupled \acs{RCP}8.5 simulations
(\range{2091}{2100}) and the 10-year mean \ac{GPP} at the end of the \nth{20}
century (\range{1991}{2000}) in the emission-driven historical simulation:
\begin{equation}
  f = \frac{\mathtxt{GPP (\range{2091}{2100})}}{
    \mathtxt{GPP (\range{1991}{2000})}} - 1.
  \label{eq:06:f}
\end{equation}
This variable is given as a percentage, where negative values correspond to a
decrease in \ac{GPP}, positive values to an increase in \ac{GPP}, and a value
of $0 \unit{\%}$ to no \ac{GPP} change. The \ac{CO2} amplitude sensitivity is
calculated for the years \range{1979}{2019}, which is the full available time
range for the observed \ac{CO2} record at \ac{KUM} \autocite{Keeling2005}.
Similar \textcite{Wenzel2016}, we extract the observed \ac{CO2} amplitude from
smoothed atmospheric \ac{CO2} concentrations (using a stiff cubic spline
function plus a four-harmonics functions with linear gain) as provided by
\textcite{Keeling2005}. For the \acs{CMIP}5 models, the respective grid cell
closest to \ac{KUM} ($19.5 \unit{\degree N}$, $154.8 \unit{\degree W}$) is
extracted. Additionally, the emission-driven historical simulations are
extended with the emission-driven \acs{RCP}8.5 simulations for the years 2006
to 2019. More details including an illustration of this calculation are shown
in \cref{fig:app:b:co2}.

Since the global constraint above cannot be used directly as a predictor in the
\ac{ML} approach which operates on gridded values (see
\cref{subsec:06:step_2}), we use it as first step to globally rescale the
\ac{ESM} output. This important first step allows us to constrain the large
uncertainty in the global \ac{CO2} fertilization effect of the \acs{CMIP}5
models. In this study, two different gridded target variables are examined (see
\cref{fig:06:schematic_steps}): the monthly climatology of future absolute
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario (see step 2a) and the fractional \acs{GPP} change over
the \nth{21} century as defined in \cref{eq:06:f} (see step 2b). The global
rescaling of the gridded target variables works as follows: Let $f_m$ be the
global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change,
and $y_{mi}$ the target variable of climate model $m$. Then, we define the
rescaled target variable $y'_{mi}$ by
\begin{equation}
  y'_{mi} = y_{mi} \cdot \frac{f'}{f_m},
  \label{eq:06:y}
\end{equation}
where the index $i$ counts over all grid cells (and months when absolute
\ac{GPP} is used as target variable in step 2a) and $m$ refers to the \ac{ESM}
($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
grid cells/months are rescaled by a factor that is determined by the ratio
between the models' global mean \ac{GPP} change and the constrained global mean
\ac{GPP} change derived from the \textcite{Wenzel2016} emergent constraint
using the observed change in the \ac{CO2} cycle amplitude. Models that simulate
a too high global mean \ac{GPP} change are nudged toward smaller values and
vice versa. The rescaled target variables are then fed into the \ac{ML} model
(step 2) presented in the following \lcnamecref{subsec:06:step_2}. We note that
this global rescaling assumes that the relative biases in the fractional
\ac{GPP} change of the individual \acp{ESM} are constant over the whole globe.

@@

For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
function $F$ relating the target variable $y$ to a set of $K$ predictors
$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (also called
\emph{features} or \emph{covariates}) such that the predicted target variable
is given by $\hat{y} = F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001,
  Friedman2002}. Its basic elements are decision trees, which use binary splits
of the input data to create decision rules. Due to their simple nature,
decision trees are easy to interpret but cannot be used to create satisfying
predictions for very complex datasets. This issue can be overcome by a
technique called \emph{boosting} \autocite{Freund1996}. Boosting iteratively
improves the performance of \emph{weak learners} (in our case decision trees)
in the following way: At first, a single decision tree is fitted to a randomly
selected subsample of the training data $\left\{ \left( \bm{x}_i, y_i \right)
\mid i \in I \right\}$ with an arbitrary index set $I$ (using a predefined loss
function, \eg{} the mean square error). Then, further decision trees are
linearly added to the regression function $F$ (scaled by a regularization
parameter $\nu \ll 1$ called \emph{learning rate}) and the model is fitted
again in the same manner. This ensures that the prediction for poorly modeled
training points in the beginning will gradually improve during the training.
The whole procedure is repeated until the loss function does not improve
anymore, as evaluated on an independent validation dataset. Relevant
hyperparameters of this algorithm (maximum number of decision trees, maximum
depth of the decision trees, learning rate, and the subsample ratio) are
optimized by $5$-fold \acf{CV} using the training data. More details on the
\ac{GBRT} algorithm are given in \cref{subsec:app:b:gbrt}.

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=\LargerFigureWidth{}]{ch06_paper_gpp/figs/2.pdf}
  \caption[
    Schematic illustration of our \acl{ML} approach to constrain projected
    absolute \acf{GPP} at the end of the \nth{21} century in step 2a.
  ]{
    Schematic illustration of our \acf{ML} approach to constrain projected
    absolute \acf{GPP} in step 2a. (a) In the training phase of the algorithm,
    the model is fitted to the training data interpolating the empirical
    (non-linear) relationship between two process-oriented diagnostics of the
    past climate $\left\{ x^{(1)}, x^{(2)} \right\}$ and (rescaled) future
    \acs{GPP} (gray surface). The points show the training points for the
    supervised \acs{ML} algorithm, each of them representing a single grid
    cell/month of a single climate model (the different colors correspond to
    different climate models). (b) In the prediction phase, observation-based
    values of the diagnostic (black points) are fed into the trained \acs{ML}
    model to constrain \acs{GPP} for every grid cell/month to a value which
    best agrees with the observations. (c) For an independent validation of our
    method, we use an out-of-sample testing setup based on a
    leave-one-model-out \acl{CV} approach (see \cref{subsec:02:model_weighting}
    for details). The schematic illustration of step 2b differs only in the
    target variable used (fractional \acs{GPP} change instead of absolute
    \acs{GPP}). \AdaptedFrom{Schlund2020}.
  }
  \label{fig:06:mlr_concept}
\end{figure}

\begin{table}[t]
  \centering
  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
      p{0.3\columnwidth}}
    \toprule
    Predictor & \makecell{Observation-driven \\ data} & Used time range &
    Physical connection to \acs{GPP} \\
    \midrule
    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
    \range{1991}{2000} & -- \\
    \midrule
    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
    grid cell. The photosynthesis rate is highly dependent on the number of
    leaves (and vegetation in general). \\
    \midrule
    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & Water is essential for the chemical processes of
    photosynthesis. \\
    \midrule
    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
    \range{1979}{2005} & Solar radiation is essential for the chemical
    processes of photosynthesis. \\
    \midrule
    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
    (incoming solar radiation). \\
    \bottomrule
  \end{tabular}
  \caption[
    Process-oriented diagnostics used in the \acf{GBRT} model to predict the
    target variables.
  ]{
    Process-oriented diagnostics (also called \emph{predictors} or
    \emph{features}) used in the \acf{GBRT} model to predict the target
    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
    variables are monthly climatologies of the specified time ranges in the
    historical climate. For step 2b (target variable: fractional \acs{GPP}
    change), the temporal mean over the specified time ranges is calculated for
    all variables. \AdaptedFrom{Schlund2020}.
  }
  \label{tab:06:predictors}
\end{table}

As already described in the previous \lcnamecref{subsec:06:step_1}, we analyze
two different target variables in this study: the monthly climatology of
absolute \ac{GPP} at the end of the \nth{21} century (step 2a) and the
fractional \ac{GPP} change over the \nth{21} century (step 2b). A schematic
representation of step 2a is shown in \cref{fig:06:mlr_concept}. The target
variable $y$ for the \ac{ML} model is the rescaled monthly climatology of
future GPP at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario given by \cref{eq:06:y}. We relate this target variable
to several process-oriented diagnostics $\bm{x}$ of the past climate
(historical simulation) listed in \cref{tab:06:predictors}. A single training
point $\left( \bm{x}_i, y_i \right)$ for the supervised \ac{ML} algorithm
corresponds to the set of the aforementioned diagnostics evaluated on a single
grid cell/month of a single climate model (see colored points in
\namecrefs{fig:06:mlr_concept} \labelcref{fig:06:mlr_concept}{\color{BrownRed}
  a} and \labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). Step 2b differs
from step 2a only in the target variable used. Both steps are independent
approaches and are evaluated separately.

@@

The raw \acs{CMIP}5 and observation-driven datasets are preprocessed before
entering the \ac{ML} algorithm (see \cref{subsec:app:b:data_preprocessing}). In
particular, increases greater than $300 \unit{\%}$ in the fractional \ac{GPP}
change that is used as target variable in step 2b are removed. These
unrealistically high values (making up about $5.9 \unit{\%}$ of the data) occur
in places with very small absolute \ac{GPP} values in the historical simulation
that are used in the denominator of the derivation of the fractional \ac{GPP}
change (see \cref{eq:06:f}). In addition, we randomly select $25 \unit{\%}$ of
the training data and use it as hold-out test data. This dataset is excluded
from the training and can therefore be used for an independent validation of
our statistical model.

@@

The \ac{SPE} of the \ac{GBRT} model can be understood as combined error due to
the internal variability of the climate system, the different model responses
within the climate model ensemble to forcing, and the incomplete description of
the climate system with the limited number of variables used as input features.
Second, the uncertainty in the rescaling is simply assessed as the uncertainty
given by the emergent constraint from step 1. Finally, the uncertainty in the
observation-based data (\ie{} uncertainty in the position of the black points
in \cref{fig:06:mlr_concept}{\color{BrownRed} b}) is assessed by error
propagation. The necessary sensitivities of the target variable to the various
predictors for each grid cell/month are given by the linear coefficients $b_k$
given by the \ac{LIME} approach introduced in \cref{subsec:06:step_2}. All
sources of uncertainty are assumed independent and are combined by adding the
squared errors, yielding a single estimate of the \ac{SPE} for each grid
cell/month. Further details on the calculation of errors are given in
\cref{subsec:app:b:uncertainty}.

@@

Using the emergent relationship, the global mean \ac{GPP} change over the
\nth{21} century can be constrained to $\pmrangeunit{39}{7}{\%}$ (standard
error), which is at the lower end of the original \acs{CMIP}5 range
($\rangeunit{31}{57}{\%}$). Moreover, the best estimate $39 \unit{\%}$ is also
slightly lower than the \acs{CMIP}5 \ac{MMM} of $43 \unit{\%}$. The observed
\ac{GPP} averaged over the years \range{1991}{2000} from the FLUXNET-MTE
product is $\pmrangeunit{123}{6}{\GtCyr{}}$ \autocite{Jung2011}. Assuming
independent errors and using Gaussian error propagation, the constrained
fractional change corresponds to a global mean \ac{GPP} of
$\pmrangeunit{171}{12}{\GtCyr{}}$ (standard error) at the end of the \nth{21}
century (\range{2091}{2100}) in the \acs{RCP}8.5 scenario. In contrast to that,
the unconstrained\acs{CMIP}5 ensemble has a model range of
$\rangeunit{156}{247}{\GtCyr{}}$ for future \ac{GPP} in the \acs{RCP}8.5
scenario (\range{2091}{2100}). The resulting global mean \ac{GPP} change is
used to rescale the gridded output of the different climate models with
\cref{eq:06:y}. The following two \lcnamecrefs{subsec:06:results_step2a} show
the results for the rescaled absolute future \ac{GPP} (step 2a) and the
rescaled fractional \ac{GPP} change (step 2b) further constrained by the
\ac{GBRT} model.

@@

\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \raisebox{5.5mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/4a.pdf}}
    \caption{}
    \label{fig:06:step2a_ml_inference:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/4b.pdf}
    \caption{}
    \label{fig:06:step2a_ml_inference:b}
  \end{subfigure}
  \caption[
    Prediction error in a leave-one-model-out \acl{CV} setup and feature
    importance of our \acl{ML} approach to constrain projected absolute
    \acf{GPP} at the end of the \nth{21} century in step 2a.
  ]{
    (a) Box plot of the \acf{RMSEP} distributions for six different statistical
    models used to predict future absolute \acf{GPP} in step 2a using a
    leave-one-model-out \acl{CV} approach. The distribution for each
    statistical model contains seven elements (black points; one for each
    climate model used as truth) and is represented in the following way: The
    lower and upper limit of the blue boxes correspond to the $25 \unit{\%}$
    and $75 \unit{\%}$ quantiles, respectively. The central line in the box
    shows the median and the black \enquote{x} shows the mean of the
    distribution. The whiskers outside the box represent the range of the
    distribution. Compared to the \acs{CMIP}5 \acf{MMM} and its corresponding
    rescaled version (r\acs{MMM}), the prediction uncertainty measured by the
    mean \acs{RMSEP} is significantly reduced by up to $48 \unit{\%}$ and $39
    \unit{\%}$, respectively, when using other statistical models. Moreover,
    the non-linear \acf{GBRT} models can slightly reduce the mean \acs{RMSEP}
    compared to the linear \acf{LASSO} models by about $2 \unit{\%}$ for
    \acs{GBRT}-1D (using historical \acs{GPP} as single predictor) and $3
    \unit{\%}$ for the full \acs{GBRT} model (using all predictors). (b)
    Relative global feature importance for the different predictors used in the
    \acs{GBRT} model to predict future absolute \acs{GPP} (step 2a). The red
    bars correspond to positive Pearson correlation coefficients between all
    predictors and the target variable. Due to its strong positive linear
    relationship with the future \acs{GPP}, the historical \acs{GPP} is by far
    the most important predictor in the model. A local feature importance map
    using \acf{LIME} is not shown here because \acs{GPP} is the dominant
    predictor for all grid cells. \AdaptedFrom{Schlund2020}.
  }
  \label{fig:06:step2a_ml_inference}
\end{figure}

>>>>

Here we briefly describe the data and methods used in this study.
\Cref{subsec:06:step_1} reviews the method of a previously published emergent
constraint on \ac{CO2} fertilization from \textcite{Wenzel2016} that we apply
here to our study to rescale the climate model output (Step 1). This includes a
discussion of the caveats that are associated with this emergent constraint
that attributes changes in the observed increase in the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} entirely to the increase in atmospheric \ac{CO2}
concentration, yet other studies provide indications that part of this increase
is due to climate and land use changes \autocite{Bastos2019, Forkel2016,
  Piao2018, Zhao2016}. In \cref{subsec:06:step_2} the \ac{ML} technique used in
Step 2 to constrain gridded \ac{GPP} projections is presented.
\Cref{fig:06:schematic_steps} shows an overview of our two-step approach.
\Cref{subsec:06:uncertainty_and_skill} describes how uncertainty and predictive
skill are quantified. A complete list of all \acs{CMIP}5 models used in this
study is given in \cref{tab:app:b:cmip5_models}. To apply the emergent
constraint in Step 1, we need emission-driven \acs{CMIP}5 simulations. The
number of \acp{ESM} included is therefore seven, as in \textcite{Wenzel2016}.
More details on the methods and the experimental setup are given in
\cref{sec:app:si_for_paper_gpp}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{ch06_paper_gpp/figs/1.pdf}
  \caption[
    Schematic illustration of our two-step approach.
  ]{
    Schematic illustration of our two-step approach. In Step 1 an emergent
    constraint by \textcite{Wenzel2016} is used to constrain the global mean
    fractional change in \acf{GPP} over the \nth{21} century. Moreover, this
    constraint is used to rescale two different gridded target variables:
    absolute \acs{GPP} at the end of the \nth{21} century and fractional
    \acs{GPP} change over the \nth{21} century. In Step 2, a \acl{ML} model is
    used to constrain these two target variables (Step 2a: absolute \acs{GPP};
    Step 2b: fractional \acs{GPP} change) in two independent approaches.
    \AdaptedFrom{Schlund2020}.
  }
  \label{fig:06:schematic_steps}
\end{figure}

@@

Therefore, our target variable in Step 1 is the global mean fractional
\acs{GPP} change over the \nth{21} century $f$ in the \acs{RCP}8.5 scenario,
which is calculated from the 10-year mean \ac{GPP} at the end of the \nth{21}
century in the emission-driven fully coupled \acs{RCP}8.5 simulations
(\range{2091}{2100}) and the 10-year mean \ac{GPP} at the end of the \nth{20}
century (\range{1991}{2000}) in the emission-driven historical simulation:
\begin{equation}
  f = \frac{\mathtxt{GPP (\range{2091}{2100})}}{
    \mathtxt{GPP (\range{1991}{2000})}} - 1.
  \label{eq:06:f}
\end{equation}
This variable is given as a percentage, where negative values correspond to a
decrease in \ac{GPP}, positive values to an increase in \ac{GPP}, and a value
of $0 \unit{\%}$ to no \ac{GPP} change. The \ac{CO2} amplitude sensitivity is
calculated for the years \range{1979}{2019}, which is the full available time
range for the observed \ac{CO2} record at \ac{KUM} \autocite{Keeling2005}.
Similar \textcite{Wenzel2016}, we extract the observed \ac{CO2} amplitude from
smoothed atmospheric \ac{CO2} concentrations (using a stiff cubic spline
function plus a four-harmonics functions with linear gain) as provided by
\textcite{Keeling2005}. For the \acs{CMIP}5 models, the respective grid cell
closest to \ac{KUM} ($19.5 \unit{\degree N}$, $154.8 \unit{\degree W}$) is
extracted. Additionally, the emission-driven historical simulations are
extended with the emission-driven \acs{RCP}8.5 simulations for the years 2006
to 2019. More details including an illustration of this calculation are shown
in \cref{fig:app:b:co2}.

Since the global constraint above cannot be used directly as a predictor in the
\ac{ML} approach which operates on gridded values (see
\cref{subsec:06:step_2}), we use it as first step to globally rescale the
\ac{ESM} output. This important first step allows us to constrain the large
uncertainty in the global \ac{CO2} fertilization effect of the \acs{CMIP}5
models. In this study, two different gridded target variables are examined (see
\cref{fig:06:schematic_steps}): the monthly climatology of future absolute
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario (see Step 2a) and the fractional \acs{GPP} change over
the \nth{21} century as defined in \cref{eq:06:f} (see Step 2b). The global
rescaling of the gridded target variables works as follows: Let $f_m$ be the
global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change,
and $y_{mi}$ the target variable of climate model $m$. Then, we define the
rescaled target variable $y'_{mi}$ by
\begin{equation}
  y'_{mi} = y_{mi} \cdot \frac{f'}{f_m},
  \label{eq:06:y}
\end{equation}
where the index $i$ counts over all grid cells (and months when absolute
\ac{GPP} is used as target variable in Step 2a) and $m$ refers to the \ac{ESM}
($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
grid cells/months are rescaled by a factor that is determined by the ratio
between the models' global mean \ac{GPP} change and the constrained global mean
\ac{GPP} change derived from the \textcite{Wenzel2016} emergent constraint
using the observed change in the \ac{CO2} cycle amplitude. Models that simulate
a too high global mean \ac{GPP} change are nudged toward smaller values and
vice versa. The rescaled target variables are then fed into the \ac{ML} model
(Step 2) presented in the following \lcnamecref{subsec:06:step_2}. We note that
this global rescaling assumes that the relative biases in the fractional
\ac{GPP} change of the individual \acp{ESM} are constant over the whole globe.

@@

For Step 2 we use the \ac{GBRT} algorithm to find an optimal regression
function $F$ relating the target variable $y$ to a set of $K$ predictors
$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (also called
\emph{features} or \emph{covariates}) such that the predicted target variable
is given by $\hat{y} = F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001,
  Friedman2002}. Its basic elements are decision trees, which use binary splits
of the input data to create decision rules. Due to their simple nature,
decision trees are easy to interpret but cannot be used to create satisfying
predictions for very complex datasets. This issue can be overcome by a
technique called \emph{boosting} \autocite{Freund1996}. Boosting iteratively
improves the performance of \emph{weak learners} (in our case decision trees)
in the following way: At first, a single decision tree is fitted to a randomly
selected subsample of the training data $\left\{ \left( \bm{x}_i, y_i \right)
\mid i \in I \right\}$ with an arbitrary index set $I$ (using a predefined loss
function, \eg{} the mean square error). Then, further decision trees are
linearly added to the regression function $F$ (scaled by a regularization
parameter $\nu \ll 1$ called \emph{learning rate}) and the model is fitted
again in the same manner. This ensures that the prediction for poorly modeled
training points in the beginning will gradually improve during the training.
The whole procedure is repeated until the loss function does not improve
anymore, as evaluated on an independent validation dataset. Relevant
hyperparameters of this algorithm (maximum number of decision trees, maximum
depth of the decision trees, learning rate, and the subsample ratio) are
optimized by $5$-fold \acf{CV} using the training data. More details on the
\ac{GBRT} algorithm are given in \cref{subsec:app:b:gbrt}.

@@

\begin{figure}[p]
  \centering
  \includegraphics[width=\LargerFigureWidth{}]{ch06_paper_gpp/figs/2.pdf}
  \caption[
    Schematic illustration of our \acl{ML} approach to constrain projected
    absolute \acf{GPP} at the end of the \nth{21} century in Step 2a.
  ]{
    Schematic illustration of our \acf{ML} approach to constrain projected
    absolute \acf{GPP} in Step 2a. (a) In the training phase of the algorithm,
    the model is fitted to the training data interpolating the empirical
    (non-linear) relationship between two process-oriented diagnostics of the
    past climate $\left\{ x^{(1)}, x^{(2)} \right\}$ and (rescaled) future
    \acs{GPP} (gray surface). The points show the training points for the
    supervised \acs{ML} algorithm, each of them representing a single grid
    cell/month of a single climate model (the different colors correspond to
    different climate models). (b) In the prediction phase, observation-based
    values of the diagnostic (black points) are fed into the trained \acs{ML}
    model to constrain \acs{GPP} for every grid cell/month to a value which
    best agrees with the observations. (c) For an independent validation of our
    method, we use an out-of-sample testing setup based on a
    leave-one-model-out \acl{CV} approach (see \cref{subsec:02:model_weighting}
    for details). The schematic illustration of Step 2b differs only in the
    target variable used (fractional \acs{GPP} change instead of absolute
    \acs{GPP}). \AdaptedFrom{Schlund2020}.
  }
  \label{fig:06:mlr_concept}
\end{figure}

\begin{table}[t]
  \centering
  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
      p{0.3\columnwidth}}
    \toprule
    Predictor & \makecell{Observation-driven \\ data} & Used time range &
    Physical connection to \acs{GPP} \\
    \midrule
    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
    \range{1991}{2000} & -- \\
    \midrule
    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
    grid cell. The photosynthesis rate is highly dependent on the number of
    leaves (and vegetation in general). \\
    \midrule
    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & Water is essential for the chemical processes of
    photosynthesis. \\
    \midrule
    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
    \range{1979}{2005} & Solar radiation is essential for the chemical
    processes of photosynthesis. \\
    \midrule
    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
    (incoming solar radiation). \\
    \bottomrule
  \end{tabular}
  \caption[
    Process-oriented diagnostics used in the \acf{GBRT} model to predict the
    target variables.
  ]{
    Process-oriented diagnostics (also called \emph{predictors} or
    \emph{features}) used in the \acf{GBRT} model to predict the target
    variables. For Step 2a (target variable: absolute \acs{GPP}), all listed
    variables are monthly climatologies of the specified time ranges in the
    historical climate. For Step 2b (target variable: fractional \acs{GPP}
    change), the temporal mean over the specified time ranges is calculated for
    all variables. \AdaptedFrom{Schlund2020}.
  }
  \label{tab:06:predictors}
\end{table}

As already described in the previous \lcnamecref{subsec:06:step_1}, we analyze
two different target variables in this study: the monthly climatology of
absolute \ac{GPP} at the end of the \nth{21} century (Step 2a) and the
fractional \ac{GPP} change over the \nth{21} century (Step 2b). A schematic
representation of Step 2a is shown in \cref{fig:06:mlr_concept}. The target
variable $y$ for the \ac{ML} model is the rescaled monthly climatology of
future GPP at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario given by \cref{eq:06:y}. We relate this target variable
to several process-oriented diagnostics $\bm{x}$ of the past climate
(historical simulation) listed in \cref{tab:06:predictors}. A single training
point $\left( \bm{x}_i, y_i \right)$ for the supervised \ac{ML} algorithm
corresponds to the set of the aforementioned diagnostics evaluated on a single
grid cell/month of a single climate model (see colored points in
\namecrefs{fig:06:mlr_concept} \labelcref{fig:06:mlr_concept}{\color{BrownRed}
  a} and \labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). Step 2b differs
from Step 2a only in the target variable used. Both steps are independent
approaches and are evaluated separately.

@@

The raw \acs{CMIP}5 and observation-driven datasets are preprocessed before
entering the \ac{ML} algorithm (see \cref{subsec:app:b:data_preprocessing}). In
particular, increases greater than $300 \unit{\%}$ in the fractional \ac{GPP}
change that is used as target variable in Step 2b are removed. These
unrealistically high values (making up about $5.9 \unit{\%}$ of the data) occur
in places with very small absolute \ac{GPP} values in the historical simulation
that are used in the denominator of the derivation of the fractional \ac{GPP}
change (see \cref{eq:06:f}). In addition, we randomly select $25 \unit{\%}$ of
the training data and use it as hold-out test data. This dataset is excluded
from the training and can therefore be used for an independent validation of
our statistical model.

@@

The \ac{SPE} of the \ac{GBRT} model can be understood as combined error due to
the internal variability of the climate system, the different model responses
within the climate model ensemble to forcing, and the incomplete description of
the climate system with the limited number of variables used as input features.
Second, the uncertainty in the rescaling is simply assessed as the uncertainty
given by the emergent constraint from Step 1. Finally, the uncertainty in the
observation-based data (\ie{} uncertainty in the position of the black points
in \cref{fig:06:mlr_concept}{\color{BrownRed} b}) is assessed by error
propagation. The necessary sensitivities of the target variable to the various
predictors for each grid cell/month are given by the linear coefficients $b_k$
given by the \ac{LIME} approach introduced in \cref{subsec:06:step_2}. All
sources of uncertainty are assumed independent and are combined by adding the
squared errors, yielding a single estimate of the \ac{SPE} for each grid
cell/month. Further details on the calculation of errors are given in
\cref{subsec:app:b:uncertainty}.

@@

Using the emergent relationship, the global mean \ac{GPP} change over the
\nth{21} century can be constrained to $\pmrangeunit{39}{7}{\%}$ (standard
error), which is at the lower end of the original \acs{CMIP}5 range
($\rangeunit{31}{57}{\%}$). Moreover, the best estimate $39 \unit{\%}$ is also
slightly lower than the \acs{CMIP}5 \ac{MMM} of $43 \unit{\%}$. The observed
\ac{GPP} averaged over the years \range{1991}{2000} from the FLUXNET-MTE
product is $\pmrangeunit{123}{6}{\GtCyr{}}$ \autocite{Jung2011}. Assuming
independent errors and using Gaussian error propagation, the constrained
fractional change corresponds to a global mean \ac{GPP} of
$\pmrangeunit{171}{12}{\GtCyr{}}$ (standard error) at the end of the \nth{21}
century (\range{2091}{2100}) in the \acs{RCP}8.5 scenario. In contrast to that,
the unconstrained\acs{CMIP}5 ensemble has a model range of
$\rangeunit{156}{247}{\GtCyr{}}$ for future \ac{GPP} in the \acs{RCP}8.5
scenario (\range{2091}{2100}). The resulting global mean \ac{GPP} change is
used to rescale the gridded output of the different climate models with
\cref{eq:06:y}. The following two \lcnamecrefs{subsec:06:results_step2a} show
the results for the rescaled absolute future \ac{GPP} (Step 2a) and the
rescaled fractional \ac{GPP} change (Step 2b) further constrained by the
\ac{GBRT} model.

@@

\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \raisebox{5.5mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/4a.pdf}}
    \caption{}
    \label{fig:06:step2a_ml_inference:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/4b.pdf}
    \caption{}
    \label{fig:06:step2a_ml_inference:b}
  \end{subfigure}
  \caption[
    Prediction error in a leave-one-model-out \acl{CV} setup and feature
    importance of our \acl{ML} approach to constrain projected absolute
    \acf{GPP} at the end of the \nth{21} century in Step 2a.
  ]{
    (a) Box plot of the \acf{RMSEP} distributions for six different statistical
    models used to predict future absolute \acf{GPP} in Step 2a using a
    leave-one-model-out \acl{CV} approach. The distribution for each
    statistical model contains seven elements (black points; one for each
    climate model used as truth) and is represented in the following way: The
    lower and upper limit of the blue boxes correspond to the $25 \unit{\%}$
    and $75 \unit{\%}$ quantiles, respectively. The central line in the box
    shows the median and the black \enquote{x} shows the mean of the
    distribution. The whiskers outside the box represent the range of the
    distribution. Compared to the \acs{CMIP}5 \acf{MMM} and its corresponding
    rescaled version (r\acs{MMM}), the prediction uncertainty measured by the
    mean \acs{RMSEP} is significantly reduced by up to $48 \unit{\%}$ and $39
    \unit{\%}$, respectively, when using other statistical models. Moreover,
    the non-linear \acf{GBRT} models can slightly reduce the mean \acs{RMSEP}
    compared to the linear \acf{LASSO} models by about $2 \unit{\%}$ for
    \acs{GBRT}-1D (using historical \acs{GPP} as single predictor) and $3
    \unit{\%}$ for the full \acs{GBRT} model (using all predictors). (b)
    Relative global feature importance for the different predictors used in the
    \acs{GBRT} model to predict future absolute \acs{GPP} (Step 2a). The red
    bars correspond to positive Pearson correlation coefficients between all
    predictors and the target variable. Due to its strong positive linear
    relationship with the future \acs{GPP}, the historical \acs{GPP} is by far
    the most important predictor in the model. A local feature importance map
    using \acf{LIME} is not shown here because \acs{GPP} is the dominant
    predictor for all grid cells. \AdaptedFrom{Schlund2020}.
  }
  \label{fig:06:step2a_ml_inference}
\end{figure}
